---
title: "ez_functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ez_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(tidyverse)
library(huggingfaceR)
```

## Introduction to EZ Functions

huggingface models are trained to perform one of several tasks. For a list of tasks, you can run `hf_list_tasks()`.

```{r huggingface_tasks_table, echo=FALSE}

tasks <- hf_list_tasks()

tasks_len <- length(tasks)

next_multiple_of_3 <- tasks_len + (0:2)[(tasks_len + 0:2) %% 3 == 0]

tasks_tbl <- tasks %>% append(rep('', next_multiple_of_3 - tasks_len)) %>%  matrix(ncol = 3, byrow = TRUE) %>% as.data.frame()

knitr::kable(tasks_tbl, row.names = NA, col.names = NULL, caption = 'Huggingface Tasks')
```

EZ functions are a simple way to get started with models for a given task. More advanced users familiar with huggingface models can load pipelines and models directly.

1.  *Beginners* --> `hf_ez_{task}`
2.  *Intermediate Users* --> `hf_load_pipeline()`
3.  *Advanced Users* --> `hf_load_tokenizer()` + `hf_load_model()` + `hf_load_AutoModel_for_task()`

So far huggingfaceR has ez functions for NLP tasks only.

```{r huggingfaceR_tasks_table, echo=FALSE}

tasks <- ls('package:huggingfaceR', pattern = "hf_ez_")

tasks_len <- length(tasks)

next_multiple_of_3 <- tasks_len + (0:2)[(tasks_len + 0:2) %% 3 == 0]

tasks_tbl <- tasks %>% append(rep('', next_multiple_of_3 - tasks_len)) %>%  matrix(ncol = 3, byrow = TRUE) %>% as.data.frame()

knitr::kable(tasks_tbl, row.names = NA, col.names = NULL, caption = 'Implemented EZ Functions')
```

## EZ Functions in Action

Let's say you'd like to automagically label the sentiment of a text as positive or negative. While you could use a text classification pre-trained model to do this, you can instead use a more generally useful zero-shot-classification model which can label text based on *any* labels you feed it. Then if you change your mind and decide that you'd like to change the labels to happy, ambivalent, and sad, you won't need to find a model explicitly trained on those sentiments!

For some sample data, we can search hugginface's datasets for a suitable set of texts.

```{r dataset}

datasets <- hf_list_datasets(pattern = "sentiment")

datasets
```

The sentiment140 dataset sounds like a good candidate. Let's load 50 rows from its training set, 25 with label 0 (negative) and 25 with label 4 (positive).

```{r}

#dataset <- hf_load_dataset(dataset = "sentiment140", split="train[0:50]")

huggingfaceR:::hf_import_datasets_transformers()

dataset <- reticulate::py$load_dataset("sentiment140", split = "train[0:25]+train[1599975:1600000]")

df <-
  dataset$data$to_pandas() %>%
  slice_sample(n = 50) %>%
  select(text, sentiment) %>%
  tibble()

knitr::kable(df, row.names = NA, col.names = NULL, caption = 'Sentiment Data')

```

We can use hf_ez_zero_shot_classification() to classify these texts as positive or negative and see how that matches up with the labels provided.

```{r}

classifier <- hf_ez_zero_shot_classification(use_api = FALSE)

zero_shot_classifications <- 
  df %>% 
  mutate(sentiment = case_when(sentiment == 0 ~ 'negative', sentiment == 4 ~ 'positive'),
         results = map(text, function(txt){
           classifier$infer(string = txt, candidate_labels = c("positive sentiment", "negative sentiment")) %>% 
             select(-sequence) %>% 
             rename_with(.fn = ~ .x %>% str_remove(" sentiment"))
         })) %>% 
  unnest(results) %>%
  mutate(match = case_when((sentiment == 'positive' & positive < negative) | (sentiment == 'negative' & negative < positive) ~ FALSE, TRUE ~ TRUE))


knitr::kable(zero_shot_classifications %>% head(10), row.names = NA, caption = 'Zero Shot Classification Results')
```

The default model used by hf_ez_zero_shot_classification(), facebook/bart-large-mnli, does a great job of classifying these texts as being positive or negative with `r zero_shot_classifications %>% count(match) %>% filter(match) %>% pull(n)` of 50 matching the given label. Impressive!
