---
title: "Getting Started with huggingfaceR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with huggingfaceR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
has_hf_token <- nzchar(Sys.getenv("HUGGING_FACE_HUB_TOKEN"))
options(cli.num_colors = 1, crayon.enabled = FALSE, pillar.subtle = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  eval = has_hf_token
)
```

```{r package-setup, include = FALSE, eval = TRUE}
library(huggingfaceR)
library(dplyr)
library(tidyr)
```

## What is huggingfaceR?

huggingfaceR provides R users with access to machine learning models hosted on
the [Hugging Face Hub](https://huggingface.co) via the Hugging Face Inference API.
You can perform natural language processing tasks -- classification, embeddings,
chat, text generation, and more -- without installing Python or managing model
weights locally. Note that the Inference API serves a curated subset of the
500,000+ models on the Hub; not every model is available for serverless inference.

Key design principles:

- **No Python required.** Authentication and a network connection are all you need.
- **Tidyverse-native.** Every function accepts character vectors and returns tibbles.
- **Pipe-friendly.** Functions compose naturally with dplyr, tidyr, and the rest
  of the tidyverse.

## Capability matrix

| Workflow | Start with | Returns |
|---|---|---|
| Sentiment and labels | `hf_classify()`, `hf_classify_zero_shot()` | one tibble row per text or label |
| Embeddings and search | `hf_embed()`, `hf_similarity()`, `hf_nearest_neighbors()` | vectors, similarities, nearest rows |
| Chat and agents | `hf_chat()`, `hf_conversation()`, `hf_tool()` | assistant messages and tool calls |
| Structured extraction | `hf_extract()` | one tidy row per input with schema columns |
| Text tasks | `hf_summarize()`, `hf_translate()`, `hf_ner()`, `hf_question_answer()` | task-specific tidy columns |
| Multimodal | `hf_transcribe()`, `hf_text_to_image()`, `hf_classify_image()` | transcripts, files/raw bytes, image labels |
| Hub workflows | `hf_hub_download()`, `hf_list_providers()`, `hf_push_dataset()` | files, provider metadata, guarded uploads |

## Installation

Install the released version from CRAN or the development version from GitHub:

```{r, eval = FALSE}
# From CRAN
install.packages("huggingfaceR")

# Development version
# install.packages("devtools")
devtools::install_github("farach/huggingfaceR")
```

## Authentication

Hugging Face requires an API token for inference requests. To obtain one:

1. Create a free account at [huggingface.co](https://huggingface.co).
2. Navigate to [Settings > Access Tokens](https://huggingface.co/settings/tokens).
3. Generate a token with at least **read** access.

Then configure the token in R:

```{r auth, eval = FALSE}
library(huggingfaceR)

# Store your token persistently (writes to .Renviron)
hf_set_token("hf_your_token_here", store = TRUE)

# Verify authentication
hf_whoami()
```

After storing the token, it is loaded automatically in future sessions.

## Quick Tour

### Classify Text

Assign labels to text using pre-trained classifiers. The default model performs
sentiment analysis, but you can supply any classification model available on the
Hugging Face Inference API. Not all models on the Hub support serverless
inference — use `hf_check_inference(model_id)` to verify.

```{r}
# Sentiment analysis
hf_classify("I love using R for data science!")

# Zero-shot classification with custom labels (no training needed)
hf_classify_zero_shot(
  "NASA launches new Mars rover",
  labels = c("science", "politics", "sports", "entertainment")
)
```

### Generate Embeddings

Convert text into dense numeric vectors that capture semantic meaning. Similar
texts produce similar vectors.

```{r}
sentences <- c(
  "The cat sat on the mat",
  "A feline rested on the rug",
  "The dog played in the park"
)

embeddings <- hf_embed(sentences)
embeddings

# Compute pairwise cosine similarity
hf_similarity(embeddings)
```

### Chat with a Language Model

Interact with open-source large language models through a simple interface.

```{r}
# Single question
hf_chat("What is the tidyverse?", max_tokens = 60)

# Guide the model with a system prompt
hf_chat(
  "Explain logistic regression in two sentences.",
  system = "You are a statistics instructor. Use plain language.",
  max_tokens = 80
)
```

### Explore the Hub

Search for models and load datasets directly into R without leaving your session.

```{r}
# Find popular text classification models
hf_search_models(task = "text-classification", limit = 5)

# Load dataset rows into a tibble
imdb <- hf_load_dataset("imdb", split = "train", limit = 100)
head(imdb)
```

### Extract Structured Data

Turn messy prose into analysis-ready columns.

```{r}
hf_extract(
  "Amelie is a chef in Paris who mentions burnout.",
  c(name = "string", occupation = "string", city = "string", theme = "string")
)
```

### Work with Images and Audio

```{r}
audio <- "https://huggingface.co/datasets/Narsil/asr_dummy/resolve/main/mlk.flac"
image <- "http://images.cocodataset.org/val2017/000000039769.jpg"

transcript <- hf_transcribe(audio, return_timestamps = "word")
substr(transcript$text, 1, 120)

hf_classify_image(image, top_k = 3)
hf_detect_objects(image, threshold = 0.5) |>
  filter(label == "cat")
```

## Working with Data Frames

All huggingfaceR functions accept character vectors and return tibbles, so they
integrate naturally into tidyverse pipelines.

```{r}
reviews <- tibble(

  product_id = 1:5,
  review = c(
    "Excellent quality, highly recommend!",
    "Broke after one week of use",
    "Good value for the price",
    "Disappointing, not as advertised",
    "Love it! Will buy again"
  )
)

# Add sentiment scores
reviews |>
  mutate(sentiment = hf_classify(review)) |>
  unnest(sentiment) |>
  select(product_id, review, label, score)
```

## Next Steps

For deeper coverage of each capability, see the following vignettes:

- [Text Classification and Zero-Shot Labeling](text-classification.html) --
  sentiment analysis, custom categories, and data frame workflows.
- [Embeddings, Similarity, and Semantic Search](embeddings-and-similarity.html) --
  vector representations, clustering, topic modeling, and visualization.
- [Chat, Conversations, and Text Generation](llm-chat-and-generation.html) --
  LLM interaction patterns, multi-turn conversations, and fill-mask.
- [Structured Extraction and Tool Calling](structured-extraction-and-tools.html) --
  JSON-schema extraction, streaming, and R-backed tool calls.
- [Multimodal: Images, Audio, and Speech](multimodal-images-audio-speech.html) --
  transcription, image generation, captioning, and object detection.
- [Hub Discovery, Datasets, and Tidymodels Integration](hub-datasets-and-modeling.html) --
  searching models, loading data, and building ML pipelines with embeddings.
- [Working with the Hub: Download, Upload, and Share](hub-download-upload-share.html) --
  file downloads, provider metadata, and guarded repository writes.
- [Analyzing the Anthropic Economic Index](anthropic-economic-index.html) --
  a research-oriented case study using embeddings, clustering, and zero-shot
  classification on real-world AI adoption data.

**For production workloads:** The classification and embeddings vignettes cover
batch processing functions (`hf_embed_batch()`, `hf_classify_batch()`, etc.)
that use parallel requests and disk checkpointing for processing thousands of
texts efficiently.

## Using Dedicated Inference Endpoints

By default, huggingfaceR sends requests to the free, serverless Hugging Face
Inference API. If you need to use a model that isn't available on the serverless
API, or you need dedicated capacity for production workloads, you can deploy a
[Dedicated Inference Endpoint](https://huggingface.co/docs/inference-endpoints)
and point huggingfaceR at it with the `endpoint_url` parameter.

```{r, eval = FALSE}
# Check whether a model supports the free serverless API
hf_check_inference("my-org/my-custom-model")

# If not, deploy a Dedicated Endpoint on huggingface.co/inference-endpoints,
# then pass its URL to any huggingfaceR function:
hf_embed(
  "Embed this with my dedicated endpoint",
  model = "my-org/my-custom-model",
  endpoint_url = "https://my-endpoint-id.us-east-1.aws.endpoints.huggingface.cloud"
)

hf_classify(
  "Classify with a private model",
  model = "my-org/my-classifier",
  endpoint_url = "https://my-endpoint-id.us-east-1.aws.endpoints.huggingface.cloud"
)

# Chat and generate also support endpoint_url
hf_chat(
  "Hello from my dedicated endpoint!",
  model = "my-org/my-llm",
  endpoint_url = "https://my-endpoint-id.us-east-1.aws.endpoints.huggingface.cloud"
)
```

The `endpoint_url` parameter is available on all inference functions, including
batch variants (`hf_embed_batch()`, `hf_classify_batch()`, etc.).

## Additional Resources

- [Hugging Face Inference API documentation](https://huggingface.co/docs/api-inference)
- [Hugging Face Inference Endpoints](https://huggingface.co/docs/inference-endpoints)
- [Hugging Face Model Hub](https://huggingface.co/models)
- [Hugging Face Datasets](https://huggingface.co/datasets)
- [Report issues on GitHub](https://github.com/farach/huggingfaceR/issues)
