---
title: "Working with the Hub: Download, Upload, and Share"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with the Hub: Download, Upload, and Share}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
has_hf_token <- nzchar(Sys.getenv("HF_TOKEN", unset = "")) ||
  nzchar(Sys.getenv("HUGGING_FACE_HUB_TOKEN", unset = ""))
options(cli.num_colors = 1, crayon.enabled = FALSE, pillar.subtle = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  eval = has_hf_token
)
```

```{r setup, eval = TRUE}
library(huggingfaceR)
library(dplyr)
```

## Inspect your token

Hub write operations require a write-scoped token. `hf_whoami()` now includes
token metadata so you can confirm the active token before creating repositories
or uploading files.

```{r, eval = FALSE}
hf_whoami() |>
  select(name, token_name, token_role, billing_mode, is_pro)
```

## Download files

Use `hf_list_repo_files()` to inspect a repository tree, then download specific
files with `hf_hub_download()`.

```{r}
files <- hf_list_repo_files("BAAI/bge-small-en-v1.5", recursive = FALSE)
files

readme <- hf_hub_download("BAAI/bge-small-en-v1.5", "README.md")
readLines(readme, n = 5)
```

Dataset and Space repositories use the same interface:

```{r}
hf_list_repo_files("stanfordnlp/imdb", repo_type = "dataset", recursive = FALSE)
```

## Search Spaces and papers

```{r}
hf_search_spaces("chat", limit = 5)
hf_search_papers("transformers", limit = 5)
```

## Choose providers before batch jobs

`hf_list_providers()` reports which Inference Providers serve a model. Coverage
comes from the Hub's provider mapping, so it works for every model and task, and
is enriched with router pricing and capability flags for chat models.

```{r}
providers <- hf_list_providers("Qwen/Qwen2.5-72B-Instruct")

providers |>
  filter(status == "live") |>
  arrange(input_price, output_price)
```

Task models such as embedding or classification models are covered too, though
they carry no router pricing metrics. Always filter on `status == "live"`: a
provider can appear in the mapping with `status == "error"` while it is
temporarily unavailable.

```{r}
hf_list_providers("BAAI/bge-small-en-v1.5")
```

`hf_check_inference()` remains a convenient one-call availability summary.

```{r}
hf_check_inference("BAAI/bge-small-en-v1.5")
```

## Guarded writes

Write operations are intentionally explicit. They require `confirm = TRUE`, a
write-scoped token, and safe defaults that avoid accidental overwrites.

```{r, eval = FALSE}
repo_id <- "your-username/analysis-results"

hf_create_repo(
  repo_id,
  repo_type = "dataset",
  private = TRUE,
  exist_ok = TRUE,
  confirm = TRUE
)

results <- tibble(
  id = 1:3,
  label = c("positive", "neutral", "negative")
)

hf_push_dataset(
  results,
  repo_id,
  path_in_repo = "results.csv",
  commit_message = "Upload analysis results",
  confirm = TRUE
)
```

Destructive deletes are more heavily guarded: `hf_delete_repo()` also refuses to
run when `CI=true`.

```{r, eval = FALSE}
hf_delete_repo(repo_id, repo_type = "dataset", confirm = TRUE)
```
