---
title: "Introduction To Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction To Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(huggingfaceR)
```

# Introduction

This vignette is the third of a three-part set which aims at two things: 
1. To get users new to Hugging Face's `transformers` library familiar with the three main abstractions:
  - Tokenizers
  - Models
  - Pipelines
2. To give users familiar with `transformers` a whistle-stop tour of the syntax they'll need to get started with each abstraction.

TODO (discuss with Sam&Alex whether this will be the recommendation)

We recommend starting with the `Introduction To Pipelines` vignette, as this gives users the quickest path to using tokenizers and models in tandem and is the highest-level abstraction.

# Models

It's important to remember when using Python that everything is an object. Models are no different, and the models that we use via `transformers` are a special type of model object - namely AutoModel objects.

As you get more comfortable using the `huggingfaceR`/ `transformers` you'll gain an appreciation for why it's important to differentiate AutoModel's from AutoModelForX's - but for now it's ok to just be aware that there is a difference.

## Loading an AutoModel

```{r}
model <- hf_load_model("bert-base-uncased")

task_model <- hf_load_AutoModel(type = "AutoModelForSequenceClassification", model_id = "bert-base-uncased")
```

