---
title: "Local embeddings and text classification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Local embeddings and text classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
run_local <- identical(Sys.getenv("HF_RUN_LOCAL_EXAMPLES"), "true")
knitr::opts_chunk$set(
  eval = run_local,
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  error = FALSE
)
options(cli.num_colors = 1, crayon.enabled = FALSE, pillar.subtle = FALSE)
```

```{r package, include=FALSE, eval=TRUE}
library(huggingfaceR)
```

## Choose where computation happens

`hf_embed()` and `hf_classify()` still use hosted inference. Nothing in this
workflow changes their defaults or initializes Python when you load the package.

The explicit local functions run embeddings and text classification on your
computer. After the initial model download, they can work offline without sending
your input text to an inference provider. There are no hosted inference charges,
but you supply the disk space, memory, and compute. Public models in this example
do not require an API token. Private or gated models may require a read token
and acceptance of the model's terms.

This first local workflow supports standard Transformers classifiers and
Sentence Transformers embedding models with **safetensors** weights. It does not
run arbitrary repository Python code, support every Hub model, or manage a local
chat server. Supported embedding module stacks use a root Transformer followed
by built-in Pooling or Normalize modules; adapter and custom-module repositories
are rejected. Without a module stack, Sentence Transformers uses mean pooling
over the local transformer. Prefer a model trained for sentence embeddings.

## Set up the optional Python stack

Install `reticulate` once if it is not already installed:

```{r install, eval=FALSE}
install.packages("reticulate")
```

```{r requirements}
hf_local_setup()
```

Setup declares the Python requirements; the first local operation may download
a compatible Python interpreter and packages, including PyTorch, Transformers,
and Sentence Transformers. This can take several minutes and substantial disk
space. The default device is the CPU.

If you already select a Python environment with `reticulate`, that choice is
respected. It must contain compatible dependencies; installing packages into a
different environment will not fix the selected one. See
[reticulate's environment documentation](https://rstudio.github.io/reticulate/articles/python_packages.html).
API-only users do not need this setup.

Use matching R and Python architectures. In particular, x64 R emulation on
Windows ARM can crash when loading Python-related R packages; successful-looking
console output does not establish that the process finished successfully.
The automated real-model runs for this workflow use native Linux CPU runners.

## Download reproducible model snapshots

Pin a full commit revision rather than relying on a moving `main` branch.
The download helper uses Hugging Face's version-aware cache and preserves the
repository's subdirectories. It retrieves safetensors weights and standard
configuration/tokenizer files, not every alternative model format.

```{r download}
cache_dir <- Sys.getenv(
  "HF_LOCAL_CACHE_DIR",
  unset = tools::R_user_dir("huggingfaceR", "cache")
)

embedding_id <- "sentence-transformers/all-MiniLM-L6-v2"
embedding_revision <- "1110a243fdf4706b3f48f1d95db1a4f5529b4d41"
classifier_id <- "distilbert/distilbert-base-uncased-finetuned-sst-2-english"
classifier_revision <- "714eb0fa89d2f80546fda750413ed43d93601a13"

embedding_path <- hf_download_model(
  embedding_id, revision = embedding_revision, cache_dir = cache_dir
)
classifier_path <- hf_download_model(
  classifier_id, revision = classifier_revision, cache_dir = cache_dir
)

data.frame(
  task = c("embed", "classify"),
  snapshot = c(basename(embedding_path), basename(classifier_path))
)
```

Downloading again reuses the cache. A snapshot is a directory containing the
model's files, not a single weight file. For example, the embedding model also
needs its pooling configuration:

```{r snapshot-layout}
list.files(embedding_path, pattern = "config.json$", recursive = TRUE)
```

## Load once, embed many texts

Loading from a downloaded directory uses local files only. Tokenizer and model
components come from the same snapshot.

```{r embedding-model}
embedding_model <- hf_load_local_model(
  embedding_path, task = "embed", local_files_only = TRUE
)
embedding_model
```

```{r embeddings}
sentences <- c(
  "The cat sat on the mat.",
  "A feline is resting on a rug.",
  "The database server needs an update."
)
embeddings <- hf_embed_local(
  sentences, embedding_model, batch_size = 2L, normalize = TRUE
)
embeddings

# A small view of the actual numeric output
round(do.call(rbind, embeddings$embedding)[, 1:8], 4)
```

The output has the same `text`, `embedding`, and `n_dims` columns as `hf_embed()`.
Each embedding is a numeric vector in a list-column.

```{r similarity}
similarities <- hf_similarity(embeddings)
similarities

stopifnot(
  all(embeddings$n_dims == 384L),
  similarities$similarity[1] > similarities$similarity[2]
)
```

These values describe this model's representation of these sentences. They are
not a general accuracy benchmark. Other embedding models may require query or
document prefixes; follow the model card when using them for retrieval.

## Classify text locally

```{r classification}
classifier <- hf_load_local_model(
  classifier_path, task = "classify", local_files_only = TRUE
)
reviews <- c(
  "I loved this movie. The acting was wonderful!",
  "I hated this movie. It was a complete waste of time."
)
classification <- hf_classify_local(reviews, classifier, batch_size = 2L)
classification

stopifnot(
  identical(classification$label, c("POSITIVE", "NEGATIVE")),
  all(classification$score > 0.9)
)
```

This English movie-review sentiment model returns one highest-scoring label per
text, with the same `text`, `label`, and `score` columns as `hf_classify()`.
Scores are model outputs, not a guarantee of calibrated confidence or suitability
for a new domain.

Long inputs are truncated to the classifier's tokenizer limit by default.
Set `truncation = FALSE` to disable that behavior; an over-length input can then
fail rather than being silently shortened. These helpers do not split long
documents into meaning-preserving chunks.

## Missing values, repeated text, and reuse

```{r missing-values}
hf_classify_local(c(reviews[1], NA_character_, reviews[1]), classifier)
hf_embed_local(c(sentences[1], NA_character_, sentences[1]), embedding_model)
```

Input order and duplicates are preserved. Missing text produces missing output,
and empty or entirely missing vectors do not invoke the prediction backend.
Classification also treats empty strings as missing text. Embedding models
accept empty strings and produce their model-specific representation.

Keep a loaded model handle for repeated calls in the same R session. A handle
contains Python objects: **do not use `saveRDS()` to move it between sessions**.
Save the model ID and revision, or the snapshot directory, and load a new handle
in the next session.

## Work offline in another R session

Once both the Python dependencies and model snapshots are present, start a new R
session with the Hub offline flag set before initializing Python:

```{r offline-example, eval=FALSE}
Sys.setenv(HF_HUB_OFFLINE = "1", TRANSFORMERS_OFFLINE = "1")
library(huggingfaceR)

cache_dir <- Sys.getenv(
  "HF_LOCAL_CACHE_DIR",
  unset = tools::R_user_dir("huggingfaceR", "cache")
)
path <- hf_download_model(
  "sentence-transformers/all-MiniLM-L6-v2",
  revision = "1110a243fdf4706b3f48f1d95db1a4f5529b4d41",
  cache_dir = cache_dir,
  local_files_only = TRUE
)
model <- hf_load_local_model(path, task = "embed", local_files_only = TRUE)
hf_embed_local("This text stays on this computer.", model)
```

An incomplete or missing cache fails explicitly; local prediction never falls
back to paid hosted inference. Cached files must remain available at the same
path. Offline model loading does not install missing Python dependencies.

## Batch size and memory

`batch_size` controls the model's inference batches, not the size of the complete
result returned to R. Embedding output is first transferred as a matrix and then
placed in a tibble list-column. At 100,000 texts and 384 dimensions, each
double-precision representation needs about 293 MiB, before model memory and
other allocations.

For larger corpora, reuse the loaded handle but process smaller R-side chunks.
Write each result to disk or pass it to a downstream consumer instead of
retaining every embedding in memory. This small example only records row counts:

```{r r-side-chunks}
indices <- split(seq_along(sentences), ceiling(seq_along(sentences) / 2L))
vapply(indices, function(index) {
  batch <- hf_embed_local(sentences[index], embedding_model)
  # Process or persist this batch here before moving to the next chunk.
  nrow(batch)
}, integer(1))
```

## Profiling the local workflow

The developer profiler compares an installed baseline with the refactored
package on the same runner, using the same cached weights, Python environment,
and two PyTorch CPU threads. It warms the models before measurement, records
seven inference repetitions for batches of 1, 32, and 256 texts, and checks that
the public API, prediction code, and output values agree.

The Python-native timing excludes transfer to R and tidy result construction.
R allocation measurements do not include Python/PyTorch memory and are **not**
peak process memory. Shared-runner timings are descriptive, not a promise of
performance on another machine or evidence of a statistically established
speedup. The profiling artifacts also contain R-only result-shaping measurements
and `Rprof()` call summaries.

```{r profile-results}
profile_dir <- Sys.getenv("HF_LOCAL_PROFILE_DIR")
if (nzchar(profile_dir)) {
  timing <- read.csv(file.path(profile_dir, "comparison.csv"))
  timing <- timing[timing$method %in% c("wrapper", "cached_hub_load"), ]
  data.frame(
    task = timing$task,
    operation = timing$method,
    texts = timing$n_texts,
    baseline_ms = round(1000 * timing$median_s_baseline, 2),
    refactored_ms = round(1000 * timing$median_s_refactored, 2)
  )
}
```

`scripts/profile-local-models.R` writes the measurements, environment metadata,
and profiles. `scripts/compare-local-profiles.R` checks the two runs and produces
the comparison table. The CI workflow pins the baseline commit explicitly;
change that reference deliberately when adopting a new performance baseline.

## Reproduce this document

Rendering is opt-in so ordinary package checks do not install Python or download
models. From an installation containing these functions, set
`HF_RUN_LOCAL_EXAMPLES=true` and render this source:

```{r render-instructions, eval=FALSE}
Sys.setenv(HF_RUN_LOCAL_EXAMPLES = "true")
rmarkdown::render("vignettes/local-models.Rmd")
```

The repository's `scripts/test-local-models.R` independently exercises downloads,
real predictions, output shapes, missing values, and cached offline reuse in a
fresh R process. Its online and offline modes also compare numeric outputs.

```{r runtime}
metadata <- reticulate::import("importlib.metadata")
python_packages <- c(
  "torch", "transformers", "sentence-transformers", "huggingface-hub", "numpy"
)
data.frame(
  component = c("R", "huggingfaceR", "reticulate", python_packages),
  version = c(
    as.character(getRversion()),
    as.character(packageVersion("huggingfaceR")),
    as.character(packageVersion("reticulate")),
    vapply(python_packages, metadata$version, character(1))
  )
)
cat(sprintf(
  "Executed: %d embeddings with %d dimensions; %d classifications.\n",
  nrow(embeddings), embeddings$n_dims[1], nrow(classification)
))
```
