---
title: "Multimodal: Images, Audio, and Speech"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multimodal: Images, Audio, and Speech}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(huggingfaceR)
library(dplyr)
```

## Transcribe interviews and podcasts

`hf_transcribe()` accepts a local audio path, URL, or raw vector and returns a
tibble with transcript text. Set `return_timestamps = "word"` when the model
supports word-level chunks.

```{r}
audio <- "https://huggingface.co/datasets/Narsil/asr_dummy/resolve/main/mlk.flac"

transcript <- hf_transcribe(audio, return_timestamps = "word")
transcript$text
transcript$chunks[[1]]
```

Transcripts are ordinary text, so you can pipe them into other package features:

```{r}
hf_extract(
  transcript$text,
  c(speaker_goal = "string", topic = "string"),
  max_tokens = 100
)
```

## Generate images

`hf_text_to_image()` writes generated images to disk and also returns the raw
bytes in a list-column.

```{r}
img <- hf_text_to_image(
  "a small red cube on a white background",
  seed = 1,
  num_inference_steps = 2,
  guidance_scale = 0
)

img$path
```

## Classify, caption, and detect objects

Image inputs can be URLs, local paths, or raw vectors.

```{r}
image <- "https://huggingface.co/datasets/huggingface/documentation-images/resolve/main/cat.png"

hf_classify_image(image, top_k = 3)
hf_caption_image(image, max_tokens = 40, temperature = 0)

boxes <- hf_detect_objects(image, threshold = 0.5)
boxes
```

Bounding-box output is numeric and ready for visualization.

```{r}
library(ggplot2)

ggplot(boxes) +
  geom_rect(aes(xmin = xmin, ymin = ymin, xmax = xmax, ymax = ymax),
            fill = NA, color = "red") +
  geom_text(aes(x = xmin, y = ymin, label = label),
            hjust = 0, vjust = 1, color = "red") +
  coord_equal()
```

## Text to speech

`hf_text_to_speech()` has the same file-output contract as
`hf_text_to_image()`, but public hosted TTS model availability varies. During
verification, beginner-friendly TTS candidates were not supported by the public
`hf-inference` provider. Use a compatible provider suffix or dedicated Inference
Endpoint when available.

```{r}
hf_text_to_speech(
  "Hello from R.",
  endpoint_url = "https://your-tts-endpoint.endpoints.huggingface.cloud"
)
```

