---
title: "Auditing GDPval Coverage and Related Hub Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auditing GDPval Coverage and Related Hub Models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 90
  chunk_output_type: console
---

```{r, include = FALSE}
env_flag <- function(name) {
  identical(tolower(Sys.getenv(name, unset = "")), "true")
}

run_public_data <- env_flag("HF_CASE_STUDY_RUN_PUBLIC_DATA")
run_inference <- env_flag("HF_CASE_STUDY_RUN_INFERENCE") &&
  (
    nzchar(Sys.getenv("HF_TOKEN", unset = "")) ||
      nzchar(Sys.getenv("HUGGING_FACE_HUB_TOKEN", unset = ""))
  )
run_geo <- env_flag("HF_CASE_STUDY_RUN_GEO")
run_heavy_embeddings <- env_flag("HF_CASE_STUDY_RUN_HEAVY_EMBEDDINGS")

options(cli.num_colors = 1, crayon.enabled = FALSE, pillar.subtle = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = run_public_data,
  message = FALSE,
  warning = FALSE
)
```

## Question and scope

[GDPval](https://huggingface.co/datasets/openai/gdpval) is a public benchmark dataset of
professional task prompts, reference files, deliverables, and rubrics. This case study
audits the dataset's coverage from R and shows how to discover Hub models that mention
GDPval in their metadata.

This article does not run benchmark evaluations or infer model capability. A model tag,
download count, or search result is metadata, not a GDPval score.

To render the live outputs, set `HF_CASE_STUDY_RUN_PUBLIC_DATA=true` before building the
article.

## Load the public Parquet file

GDPval contains list columns for reference and deliverable files. Reading the public
Parquet file preserves those columns cleanly.

```{r load-gdpval}
library(dplyr)
library(stringr)
library(tibble)

if (!requireNamespace("arrow", quietly = TRUE)) {
  stop("Install the 'arrow' package to build this article.", call. = FALSE)
}

gdpval_url <- paste0(
  "https://huggingface.co/datasets/openai/gdpval/resolve/main/",
  "data/train-00000-of-00001.parquet"
)

gdpval <- arrow::read_parquet(gdpval_url, as_data_frame = TRUE)

required_cols <- c(
  "task_id", "sector", "occupation", "prompt",
  "reference_files", "reference_file_urls", "reference_file_hf_uris",
  "deliverable_files", "deliverable_file_urls", "deliverable_file_hf_uris",
  "rubric_pretty", "rubric_json"
)
stopifnot(all(required_cols %in% names(gdpval)))

gdpval <- gdpval |>
  mutate(
    prompt_chars = nchar(prompt),
    n_reference_files = lengths(reference_files),
    n_deliverable_files = lengths(deliverable_files),
    rubric_chars = nchar(rubric_pretty)
  )

print(tibble(
  source_dataset = "openai/gdpval",
  source_file = "data/train-00000-of-00001.parquet",
  rows = nrow(gdpval),
  columns = ncol(gdpval) - 4,
  sectors = n_distinct(gdpval$sector),
  occupations = n_distinct(gdpval$occupation),
  median_prompt_chars = median(gdpval$prompt_chars),
  zero_reference_tasks = sum(gdpval$n_reference_files == 0),
  max_reference_files = max(gdpval$n_reference_files)
), width = Inf)
```

The first summary establishes the basic benchmark shape: 220 tasks, 12 source columns, 9
sectors, and 44 occupations. Prompt length and reference-file counts describe benchmark
packaging, not task difficulty or model performance.

## Inspect the source schema

```{r schema}
print(tibble(
  column = required_cols,
  present = required_cols %in% names(gdpval),
  role = c(
    "unique task identifier",
    "sector label",
    "occupation label",
    "task instructions",
    "supporting file names",
    "supporting file URLs",
    "supporting file Hugging Face URIs",
    "expected deliverable file names",
    "expected deliverable file URLs",
    "expected deliverable Hugging Face URIs",
    "human-readable rubric",
    "machine-readable rubric JSON"
  )
), n = Inf, width = Inf)
```

The deliverable and rubric fields are central to GDPval. They are the evidence needed for
an actual evaluation, so a metadata audit should confirm they are present before any model
claims are made.

## Audit sector and occupation coverage

```{r coverage}
sector_counts <- gdpval |>
  count(sector, sort = TRUE)

occupation_summary <- gdpval |>
  count(occupation, sort = TRUE) |>
  summarise(
    occupations = n(),
    min_tasks_per_occupation = min(n),
    median_tasks_per_occupation = median(n),
    max_tasks_per_occupation = max(n)
  )

print(sector_counts, n = Inf, width = Inf)
print(occupation_summary, width = Inf)
```

These counts describe benchmark coverage. They should not be read as the distribution of
activity in the broader economy or as sector-level AI exposure.

## Audit prompt and file burden

```{r prompt-file-burden}
prompt_summary <- gdpval |>
  summarise(
    min_prompt_chars = min(prompt_chars),
    median_prompt_chars = median(prompt_chars),
    max_prompt_chars = max(prompt_chars),
    median_rubric_chars = median(rubric_chars)
  )

reference_distribution <- gdpval |>
  count(n_reference_files, name = "tasks") |>
  arrange(n_reference_files)

deliverable_distribution <- gdpval |>
  count(n_deliverable_files, name = "tasks") |>
  arrange(n_deliverable_files)

print(prompt_summary, width = Inf)
print(reference_distribution, n = Inf, width = Inf)
print(deliverable_distribution, n = Inf, width = Inf)
```

The reference-file distribution shows whether a task is pure prompt reading or requires
supporting documents. That matters for evaluation logistics, but it is not by itself a
measure of task difficulty.

## Check list-column integrity

For reference and deliverable metadata, the file names, URLs, and Hugging Face URIs should
have matching lengths within each task.

```{r list-integrity}
integrity <- gdpval |>
  transmute(
    reference_lengths_match =
      lengths(reference_files) == lengths(reference_file_urls) &
      lengths(reference_files) == lengths(reference_file_hf_uris),
    deliverable_lengths_match =
      lengths(deliverable_files) == lengths(deliverable_file_urls) &
      lengths(deliverable_files) == lengths(deliverable_file_hf_uris),
    has_rubric_pretty = !is.na(rubric_pretty) & rubric_chars > 0,
    has_rubric_json = !is.na(rubric_json) & nchar(rubric_json) > 0
  )

integrity_summary <- integrity |>
  summarise(
    reference_lengths_all_match = all(reference_lengths_match),
    deliverable_lengths_all_match = all(deliverable_lengths_match),
    tasks_with_rubric_pretty = sum(has_rubric_pretty),
    tasks_with_rubric_json = sum(has_rubric_json)
  )

print(integrity_summary, width = Inf)
```

This check is a guardrail before downloading files or scoring model outputs. If file-name
and URI lengths do not align, downstream evaluation code could silently use the wrong
supporting materials.

## Inspect reference-heavy examples

```{r task-examples}
reference_heavy_examples <- gdpval |>
  arrange(desc(n_reference_files), desc(prompt_chars)) |>
  transmute(
    task_id,
    sector,
    occupation,
    prompt_chars,
    n_reference_files,
    n_deliverable_files,
    prompt_preview = str_trunc(str_squish(prompt), 90)
  ) |>
  slice_head(n = 8)

print(reference_heavy_examples, n = Inf, width = Inf)
```

These examples are useful for planning an evaluation run. Reference-heavy tasks may require
file parsing, retrieval, or multimodal handling in addition to text generation.

## Discover Hub model metadata

The Hub can be searched for model metadata that mentions GDPval. Treat these results as
leads, not benchmark scores.

```{r model-metadata}
model_mentions <- tryCatch(
  {
    huggingfaceR::hf_search_models(search = "gdpval", limit = 10) |>
      transmute(
        model_id,
        task,
        downloads,
        likes,
        evidence = "Hub model search result for 'gdpval'"
      )
  },
  error = function(e) {
    tibble(
      model_id = paste("Hub search failed:", conditionMessage(e)),
      task = NA_character_,
      downloads = NA_integer_,
      likes = NA_integer_,
      evidence = "Search did not complete"
    )
  }
)

print(model_mentions, n = Inf, width = Inf)
```

A search result means the model card or metadata references GDPval. It does not mean the
model is official, that it was fine-tuned on GDPval, or that it achieved any particular
benchmark score.

## Evidence ladder for model claims

```{r evidence-ladder}
print(tibble(
  evidence_level = c(
    "Official GDPval result",
    "Reproducible evaluation run",
    "Model card self-report",
    "Hub metadata or search hit"
  ),
  safe_wording = c(
    "The official source reports this result.",
    "This exact model/version was run on GDPval with documented prompts, files, and scoring.",
    "The model card reports a GDPval-related result; verify the method before comparing.",
    "The model metadata mentions GDPval; this is only a discovery lead."
  )
), n = Inf, width = Inf)
```

The key distinction is evidence. Metadata discovery helps find candidate models or cards to
read next, but it should not be turned into a leaderboard.

## Takeaways

This case study audited GDPval as a public benchmark dataset, not as a model-performance
claim. The reusable workflow is: load the Parquet file, verify the schema, summarize task
and file coverage, inspect rubrics and deliverables, and treat Hub model search results as
metadata leads only.
