---
title: "sentence_tranformers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sentence_tranformers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Aside from wrapping Hugging Face's `transformers` library, `huggingfaceR` also wraps around the `sentence_transformers` library to allow SOTA document embeddings. 

```{r setup}
library(huggingfaceR)
library(dplyr)
library(stringr)
```

It's incredibly simple to use `sentence_transformers` in `huggingfaceR`. You just need one function: `hf_load_sentence_model` and a model_id. To begin with, we'll use the 'paraphrase-MiniLM-L6-v2' as our model:

```{r loading a model}
model_id <- "paraphrase-MiniLM-L6-v2"

minilm_l6 <- hf_load_sentence_model(model_id)
```

The first, and for many the only, thing a user will want to do is feed in some document(s) and receive the embedding(s). We can do that by using R's `$` syntax to access our sentence transformer's `encode` class method - if unfamiliar with OOP/Python, just think of class methods as functions.

But first, we'll need a document:

```{r output}
document <- c("Many people think Lionel Messi is the greatest footballer to have ever played the Beautiful Game.")

(embedding <- minilm_l6$encode(document))
```

Calling our model on one document returns a 384 length vector. We could tidy it into a tibble by first transposing:
```{r tidying the output}
embedding %>%
  t()  %>%
  as_tibble()
```

Most likely we're not going to want to embed just one document, but many. Let's use `stringr::sentences` as a test run. There are 720 sentences which we'll embed in one go. Because we're embedding a number of sentences, we'll set `show_progress_bar = TRUE`, we'll also change the `batch_size to 64L` - although the default setting of 32L would be fine too.

```{r}
start <- proc.time()
sentences_embeddings <- minilm_l6$encode(stringr::sentences,show_progress_bar = TRUE,
                                         batch_size = 64L)
proc.time() - start


```

The process took about 2 seconds start to finish, not bad! We can tidy up our results similarly to before:

```{r}
sentences_embeddings %>%
  t() %>%
  as_tibble()
```

However, when dealing with more and longer documents, or when using a larger model, the proces can take significantly longer. In such cases, it's a good idea to use a GPU if you have access to one. The author is currently using a Macbook with an M1 chip.  There is ongoing work by the Pytorch team to accelerate MPS chips, currently we can send some models to our GPU and some not - most Hugging Face models require `aten::cumsum.out` for which the backend integration with Apple Silicon chips has not yet been written. The process ought to be similar, and simpler with an NVIDIA/Cuda GPU.

You will need to have the appropriate torch version installed for you GPU.

```{r}
reticulate::py_run_string("import torch")

device <- hf_set_device()

minilm_gpu <- hf_load_sentence_model(model_id)
minilm_gpu$to(device)
```

Now we're ready to call our model with GPU acceleration:

```{r}
sentences_embeddings_gpu <- minilm_gpu$encode(stringr::sentences, device = device, show_progress_bar = TRUE, batch_size = 64L)
```

You'll notice that it actually takes longer with the GPU - that's because of the set up costs involved. With a larger dataset or model you should expect to see at least a 3x speed up when using an M1 chip, and more with an NVIDIA/cuda gpu (depending on hardware of course).

Let's say you wanted a different model to "paraphrase-MiniLM-L6-v2". You could use `huggingfaceR::models_with_downloads` to select a model based on downloads:
```{r}
models_with_downloads %>%
  filter(task == "sentence-similarity")
```

```{r}
model_id <- "sentence-transformers/stsb-distilbert-base"
st_distilbert <- hf_load_sentence_model(model_id)
```

And you could extract embeddings using the same methods as before. Note if using RStudio - when calling model$encode() you can use tab to access available arguments. 
```{r}
st_embeddings <- st_distilbert$encode(stringr::sentences, show_progress_bar = TRUE)

```

Voila, for most use cases you should be covered.

# Advanced Users

Let's take a step back, we can use R's familiar `$` syntax to access our model's class methods and config.
```{r}
names(minilm_l6)
```

Our original model outputs 384 dimensional embeddings, and accepts up to 128 tokens:

```{r}
c(minilm_l6$get_sentence_embedding_dimension(),minilm_l6$max_seq_length)
```
 
Our next model outputs 768 dimensional embeddings and also accepts up to 128 tokens:

```{r}
c(st_distilbert$get_sentence_embedding_dimension(), st_distilbert$max_seq_length)
```

There's a lot you can do/look at - for example you can get your model's tokenizer's full vocabulary:

```{r}
model_vocab <- st_distilbert$tokenizer$vocab

model_vocab_df <- tibble(
  token = names(model_vocab),
  id = unlist(model_vocab, recursive = FALSE)
)

model_vocab_df
```

The possibilities are (almost) endless.
