---
title: "Structured Extraction and Tool Calling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structured Extraction and Tool Calling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(huggingfaceR)
library(dplyr)
```

## Why structured output matters

Chat models are useful for summarizing messy text, but analysis usually needs
columns. `hf_extract()` asks a chat model for JSON that conforms to a schema and
returns one tidy row per input.

```{r}
notes <- c(
  "Amelie is a chef in Paris who mentions burnout.",
  "Jordan is a nurse in Seattle who mentions staffing."
)

hf_extract(
  notes,
  c(name = "string", occupation = "string", city = "string", theme = "string"),
  max_tokens = 120
)
```

The lightweight schema form is enough for many workflows. Use a full JSON Schema
list when you need enums, descriptions, or nested objects.

```{r}
schema <- list(
  type = "object",
  properties = list(
    name = list(type = "string"),
    sentiment = list(type = "string", enum = c("positive", "neutral", "negative")),
    urgency = list(type = "number")
  ),
  required = c("name", "sentiment", "urgency"),
  additionalProperties = FALSE
)

hf_extract("Priya needs help today but likes the program.", schema)
```

## Tool definitions

`hf_tool()` builds OpenAI-compatible function definitions for models/providers
that support tool calling. The default chat model is chosen for easy first calls;
for tools, use a tool-capable model such as `Qwen/Qwen2.5-72B-Instruct`.

```{r}
add_tool <- hf_tool(
  "add",
  "Add two numbers.",
  c(x = "number", y = "number")
)

res <- hf_chat(
  "Use the add tool to add x=2 and y=3.",
  model = "Qwen/Qwen2.5-72B-Instruct",
  tools = list(add_tool),
  tool_choice = "auto",
  temperature = 0
)

res$tool_calls
```

## Running tools in a conversation

`hf_run_tools()` executes tool calls requested by the assistant, appends tool
result messages, and asks the model for the final answer.

```{r}
convo <- hf_conversation(model = "Qwen/Qwen2.5-72B-Instruct")
convo <- chat(
  convo,
  "Use the add tool to add x=2 and y=3, then tell me the answer.",
  tools = list(add_tool),
  tool_choice = "auto",
  temperature = 0
)

convo <- hf_run_tools(
  convo,
  list(add = function(x, y) x + y),
  temperature = 0
)

convo$history
```

## Streaming callbacks

Set `stream = TRUE` to receive deltas as they arrive. When `callback` is omitted,
deltas print to the console.

```{r}
pieces <- character()

answer <- hf_chat(
  "Reply with exactly: OK",
  stream = TRUE,
  callback = function(delta) pieces <<- c(pieces, delta),
  temperature = 0,
  max_tokens = 8
)

paste0(pieces, collapse = "")
answer
```

